---
title: "CART MODELS"
author: "Karthik"
date: "2/15/2020"
output: html_document
---

```{r}
#install.packages("rpart")
library(rpart)
#install.packages("rpart.plot")
library(rpart.plot)
library(caTools)
```


```{r}
head(CART_Train,10)

```



```{r}
sum(CART_Train$Personal.Loan=="1")
sum(CART_Test$Personal.Loan=="1")

```

```{r}
cartParameters = rpart.control(minisplit=30,minibucket=10,cp=0,xval=10)
cartModel_Train=rpart(formula=Personal.Loan~.,data=CART_Train[,c(-1,-5)],method="class",
                control=cartParameters)
cartModel_Train
```

Based on the CART model output it can be said that there 336 GOOD predictions out of 3500 datasets. This shows a unbalance nature of the dataset.Probability of good prediciton is about 9.6% while 90.4% probability of bad prediction.

```{r}
View(CART_Train)
names(CART_Train)
rpart.plot(cartModel_Train,cex = NULL,roundint=FALSE,
           tweak = 1.5,fallen.leaves = TRUE,
           varlen = -3,faclen = -3,box.palette=c("green","green2","green4","orange"))
```



```{r}
#install.packages("rattle")
library(rattle)
#install.packages("RColorBrewer")
library(RColorBrewer)
par(mfrow=c(1,1))
fancyRpartPlot(cartModel_Train)
```


```{r}
plotcp(cartModel_Train) # Plot of a relativeerrror

printcp(cartModel_Train) # Print the relative error

```


Now let's prune the tree to obtain a better CART model. The following command will calculate a best CP value.

```{r}
bestcp=cartModel_Train$cptable[which.min(cartModel_Train$cptable[,"xerror"]),"CP"]
bestcp
```

```{r}
  tree =   rpart(formula = Target ~ ., 
                 data = trainDS, method = "class", minsplit=100, minbucket = 10, cp = 0.0021, xval = 10)
```



```{r}
ptree = prune(cartModel_Train, cp= 0.00297619 ,"CP")
printcp(ptree)

ptree
rpart.plot(ptree,roundint=FALSE,tweak = 1.0,
           fallen.leaves = TRUE,varlen = -3,faclen = -3,
           box.palette=c("green","green2","green4","orange"))

```


Finally, we might want to use the decision tree to predict the 
class for each row and/or score the probabilities:
```{r}

CART_Train$predict.class = predict(ptree, CART_Train, type = "class")
CART_Train$prob_1 = predict(ptree, CART_Train, type = "prob")[,"1"]
head(CART_Train,20)
View(CART_Train)
```
As shown in the above table, it can be said that all personal loan 1 prediction have higer probability in the range of approximately 97%.










