---
title: "Mushroom Case Study - EDA"
author: "Karthik"
date: "10/13/2020"
output: html_document
---


#1.0 Working Directory
```{r}
setwd("F:/Data Science/ebuilder/")
getwd()
```

##1.1 Importing Dataset
```{r}
dataset <- read.csv("mushrooms.csv")
#mush <- read.csv("mushrooms.csv")
```

##1.2 Loading Libraries
```{r}
library("naniar") # NA plots
library("mice") # NA visualisations
library("ggplot2") # Visualising plots
library("dplyr") # Variable Transofrmation
library("klaR") # K Mode clustering techniques
library("DMwR") # KNN imputations for NA
library("cba")
library("caTools")
```
#EDA
##2.0 Dimentions Analysis
```{r}
dim(dataset)
```

```{r}
colnames(dataset)
```

## 2.1 Header Changed Dataset
```{r}

names(dataset) <- c("edibility","cap.shape","cap.surface","cap.color","bruises","odor","gill.attachment","gill.spacing","gill.size","gill.color","stalk.shape","stalk.root","stalk.surface.above.ring","stalk.surface.below.ring","stalk.color.above.ring","stalk.color.below.ring","veil.type","veil.color","ring.number","ring.type","spore.print.color","population","habitat")

```

## 2.2 Chnge Attribute Classes
```{r}
str(dataset)

dataset[,] <-lapply(dataset[,],factor) # Converting all variables classed into Factor.
```

## 2.3 Converting the Missing value "?" into NA
```{r}
summary(dataset) # 5 Number Summary.

levels(dataset$stalk.root)<- c(NA,"b","c","e","r") # "?" level into NA.
levels(dataset$stalk.root)
```


#...
##3.0 Missing Value Visualisation & Treatment

```{r}
anyNA(dataset) #  Presence of overall missing value
sum(is.na(dataset)) # Sum of overall NAs 
colSums(is.na(dataset)) # Column wise presence of NAs.
```

## 3.1 NA Visualisations
```{r}
library("visdat")
vis_dat(dataset) # Entire dataset in one View.See the class information and NA values in the dataset.

```

```{r}
vis_miss(dataset) # NA Visualisations
```
Well only 1.3% of the total data have been missing. Also, missing entry can be found only on the "stalk.root" attribute.

##3.2 Relationship of Missing Data
```{r}
gg_miss_var(dataset,facet=edibility) # NA visualisation based on edibility
```

#------------------------------------
##4.0 Data Transformation for Analysis

```{r}
# Analysis on Edibility only on observations contains NA entries.
dataset %>% filter( edibility=="p") %>% count(is.na(stalk.root))
edible <-dataset %>% filter(is.na(stalk.root) & edibility=="e") # Edible mushrooms with NA entries
dim(edible)


poison <- dataset %>% filter(is.na(stalk.root),edibility=="p") # poisonous mushrooms with NA entries
dim(poison)
(count(poison)/nrow(dataset))*100 # Poisonous mushroom percentage
(count(edible)/nrow(dataset))*100 # Edible mushroom percentage
```
Among the overall missing values 8.7% of missed stalk.root information are of Poisonous mushrooms while  21.7% of missing stalk.root information are of edible mushrooms.

##4.1  Creating Multiple datasets (with Treated & Untreated NA)
Misclassification of edibility of mushrooms could cause human fatality.Hence,I have decided to use the following dataset.
1. Complete Observations - Data 
2. NA Treated Observations -DataT
3. NA Treated on Edible Mushrooms Observations only. - DataE

```{r}
Data <- dataset[complete.cases(dataset),] # NA removed dataset
any_na(Data)
dim(Data)

DataT <- knnImputation(dataset, k = 5) # NA Treated dataset
any_na(DataT)
dim(DataT)

DataE <-knnImputation((dataset %>% filter(!(edibility=="p" & is.na(stalk.root)))), k = 5) # Edible observations NA treated dataset.
dim(DataE)
any_na(DataE)


```
#--------------------------------





##5.0 Train & Test Data Split
###5.1 Data Split - Complete Cases Only
```{r}
#Data
set.seed(369)
split=sample.split(Data$edibility,SplitRatio = 0.7)
train.D <- subset(Data,split==TRUE)
test.D <- subset(Data ,split==FALSE)

dim(train.D)
dim(test.D)
```
###5.2 DataT Split - NA Treated 
```{r}
#DataT
set.seed(369)
split=sample.split(DataT$edibility,SplitRatio = 0.7)
train.T <- subset(DataT,split==TRUE) # 70% Train Split
test.T <- subset(DataT,split==FALSE) # 30% Test Split

dim(train.T)
dim(test.T)
```

###5.3 DataE Split - NA Treated Edible only
```{r}
#DataE
set.seed(369)
split=sample.split(DataE$edibility,SplitRatio = 0.7)
train.E <- subset(DataE,split==TRUE)
test.E <- subset(DataE,split==FALSE)

dim(train.E)
dim(test.E)
```


