---
title: "IPL EDA"
author: "Karthik"
date: "11/25/2020"
output: html_document
---



# Working Directory
```{r}
setwd("F:/Data Science/Casestudies/IPL/")
getwd()
```
# Libraries
```{r}
library(readxl) # Importing Excel files into R
library(tidyverse) # Pipeline command & Data transformation
library(GGally) # GG pairs plot
library(corrplot)# Plotting Correlation Matrix
#install.packages("GGally")
library(ggplot2)# Data Visualisation
library(visdat) # Missing Value Visualisations
```


# Import Dataset
```{r}
season_bat_df <- read_xlsx("season.xlsx",sheet = "Bat") # Season Batsmen Dataset
career_bat_df <- read_xlsx("career.xlsx",sheet = "Bat") # Career Batsment Dataset
scorecards_bat_df <- read_xlsx("scorecards.xlsx",sheet = "Bat") # Scorecard Dataset
output <- read_xlsx("Output.xlsx") # Output dataset
```

#-------------------------------------------------------------------------------
# NOTES: 

I) Investigate Players overall IPL career.
        1) Pick best players based on KPI.
            a)Most Played. OK
            b)Most Batted. OK
            c)Ave,SR - Max and Min Indicates overall IPL career.
            d)Give a confidence score.
            e)Runs,BF.
            f)inn,NO.

            
II) Investigate Player by Season.
    1)Pick best player of the season.
        a) Ave,SR - Max and Min.
        b) Most Played.
        c) Most Batted.
    2)Investigate Player's best season based on best career.
    3)Investigate Player's worst season based on best career. 
    
III) In scorecard inni data dictionary - 0 =1st inn, 1= 2nd inn.
     Actual dataset 1 and 2.
     1) Most scored player against which IPL team.
     2) Best player season and career against which IPL team.
     3) Worst player season and career against which IPL team.
#-------------------------------------------------------------------------------
#2.0 EDA 
##2.1 Sanity Check  
###2.1.2 Variable Names
```{r}
career_bat <- career_bat_df # Setting dataset for processing

# Coloumn name transformation
colnames(career_bat) <- c("Player","Mat.c","Inns.c","NO.c","Runs.c","BF.c","hundreds.c","fifties.c","ducks.c","four.c","six.c","Ave.c","SR.c","HS.c","NO_HS.c" )

# Display Coloumn names
colnames(career_bat)
```

Attributes which describe same information has different names on different excel sheet. In order to create a standardize understanding in our interpretation converted column names into a common name.  

###2.1.3 Class Type

Let us investigate on the data type of the given "career" data set attributes.
```{r}
str(career_bat)# Display variable classes

# Transforming variable classes.
career_bat$Player <- as.factor(career_bat$Player)
career_bat$Inns.c <- as.integer(career_bat$Inns.c)
career_bat$Mat.c <- as.integer(career_bat$Mat.c)
career_bat$NO.c <- as.integer(career_bat$NO.c)
career_bat$BF.c <- as.integer(career_bat$BF.c)
career_bat$NO_HS.c <-as.factor(career_bat$NO_HS.c) 
career_bat$Runs.c <- as.integer(career_bat$Runs.c)
career_bat$four.c <- as.integer(career_bat$four.c)
career_bat$six.c <- as.integer(career_bat$six.c)
career_bat$HS.c <- as.integer(career_bat$HS.c)
career_bat$hundreds.c <-as.integer(career_bat$hundreds.c) 
career_bat$fifties.c <-as.integer(career_bat$fifties.c) 
career_bat$ducks.c <-as.factor(career_bat$ducks.c) 
```
## 2.2 Missing Values

The "career" dataset contains a complete rows of observation hence, we could proceed with further analysis.
```{r}
anyNA(career_bat) # Checking for missing values in Career dataset
summary(career_bat) # Generating 5 Number summary f all attributes.
dim(career_bat)
```

There are about 560 batsmen have been taken into consideration and their performance have been evaluated using  14 attributes.

#3.0 UNIVARIATE ANALYSIS
##3.1 PLAYER

```{r}
## Most Played
career_bat %>% 
  select(Player,Mat.c) %>%  
  filter(Mat.c==max(Mat.c)) # Most number of matches played by a batsman
## Least Played
career_bat %>% 
  select(Player,Mat.c) %>%  
  filter(Mat.c==min(Mat.c)) %>% 
  count() # All players have played at least one IPL match in their career. 

## Most Batted
career_bat %>% 
  select(Player,Inns.c) %>% 
  filter(Inns.c==max(Inns.c)) 


## Most Batted
career_bat %>% 
  select(Player,Inns.c) %>% 
  filter(Inns.c==min(Inns.c)) %>%count()/nrow(career_bat)*100 

Top_ten_inn <- top_n(career_bat,10,Inns.c) %>% arrange(desc(Inns.c))
Top_ten_inn    

```

As per the given data set SK Raina have played the most number of IPL matches in his career which is 193 matches. Also, interesting to note that, 45 players have played only one IPL match so far.

Evidently, he holds the first place for a batsman who batted the most number of games he played that is 99 matches. In contrast, all batsmen have got a chance to bat at least once in their career. It is 7.5% of total list.



## 3.2 Variable Transformation.
```{r}
career_bat_new<- career_bat %>% transmute(Player,
            Mat.c,
            Inns.c,
            NO.c,
            Runs.c,
            BF.c,
            hundreds.c,
            fifties.c,
            ducks.c,
            boundaries.c= four.c+six.c,
            Ave.c,
            SR.c,
            HS.c,
            NO_HS.c) %>% select(Player,Inns.c,boundaries.c,Ave.c,SR.c,NO_HS.c)
```

##3.3 Inns
"Inns" denotes the number of innings played by each batsmen.
```{r}

range(career_bat_new$Inns.c) # Min & Max of Inns played.

#Histogram
ggplot(career_bat_new,aes(x=Inns.c))+geom_histogram(color="darkblue",fill='lightblue',binwidth=1)+
  geom_vline(aes(xintercept=mean(Inns.c)),color="black",linetype="dashed",size=1)+
  geom_vline(aes(xintercept=median(Inns.c)),color ="red",linetype="dashed",size=1)+
  labs(title="Innings played Histogram Plot",x="Innings Played", y = "Count")

# Box plot    
boxplot(career_bat_new$Inns.c,horizontal = TRUE,col="blue",main = "Innings Played Boxplot Distribution", xlab= "Number of Innings played")
```
```{r}
# Innings Played Outliers

# OUTLIER CALCULATION
mean(career_bat_new$Inns.c) # Average Inns played in their career
median(career_bat_new$Inns.c) # Median of the Inns played in their career.

IQR_Inns.c = IQR(career_bat_new$Inns.c)
IQR_Inns.c

# Quantile Caluculation
Q3_Inns.c = quantile(career_bat_new$Inns.c,0.75)
Q3_Inns.c

Q1_Inns.c =quantile(career_bat_new$Inns.c,0.25)
Q1_Inns.c

# IQR Calulation
Max_IQR_Inns.c = Q3_Inns.c + 1.5*IQR_Inns.c
Min_IQR_Inns.c = Q1_Inns.c + 1.5*IQR_Inns.c

# Outliers calculation
sum(career_bat_new$Inns.c > Max_IQR_Inns.c)
sum(career_bat_new$Inns.c < Min_IQR_Inns.c)

```
Innings played in their IPL career distribution contains 68 lower outliers and 452 upper outliers.
However, on average each player have batted at least 20 games in their career. An interesting observation is, at least 25% of the players have batted 22 games.

##3.4 Boundaries Scored
This attribute is a summation of both fours and sixers scored by each batsmen.


```{r}

ggplot(career_bat_new,aes(x=boundaries.c))+geom_histogram(color="darkblue",fill='lightblue',binwidth=1)+
  geom_vline(aes(xintercept=mean(boundaries.c)),color="black",linetype="dashed",size=1)+
  geom_vline(aes(xintercept=median(boundaries.c)),color ="red",linetype="dashed",size=1)+
  labs(title="Innings played Histogram Plot",x="Innings Played", y = "Count")



```


```{r}


# Boundaries scored Outliers

# OUTLIER CALCULATION
mean(career_bat_new$boundaries.c) # Average Inns played in their career
median(career_bat_new$boundaries.c) # Median of the Inns played in their career.

IQR_boundaries.c = IQR(career_bat_new$boundaries.c)
IQR_boundaries.c

# Quantile Caluculation
Q3_boundaries.c = quantile(career_bat_new$boundaries.c,0.75)
Q3_boundaries.c

Q1_boundaries.c =quantile(career_bat_new$boundaries.c,0.25)

Q1_boundaries.c

# IQR Calulation
Max_IQR_boundaries.c = Q3_boundaries.c + 1.5*IQR_boundaries.c
Min_IQR_boundaries.c = Q1_boundaries.c + 1.5*IQR_boundaries.c

# Outliers calculation
sum(career_bat_new$boundaries.c > Max_IQR_boundaries.c)
sum(career_bat_new$boundaries.c < Min_IQR_boundaries.c)
```

Most number of boundaries played by single player in his IPL career is 694. There are 444 highly entertained IPL players who have scored the most number of boundaries and considered to be an outlier.

```{r}
# Innings Played Outliers
ggplot(career_bat_new,aes(x=SR.c))+geom_histogram(color="darkblue",fill='lightblue',binwidth=1)+
  geom_vline(aes(xintercept=mean(SR.c)),color="black",linetype="dashed",size=1)+
  geom_vline(aes(xintercept=median(SR.c)),color ="red",linetype="dashed",size=1)+
  labs(title="Innings played Histogram Plot",x="Innings Played", y = "Count")

# OUTLIER CALCULATION
mean(career_bat_new$SR.c) # Average SR.c in their career
median(career_bat_new$SR.c) # Median of the SR.c played in their career.

IQR_SR.c = IQR(career_bat_new$SR.c)
IQR_SR.c

# Quantile Caluculation
Q3_SR.c = quantile(career_bat_new$SR.c,0.75)
Q3_SR.c

Q1_SR.c =quantile(career_bat_new$SR.c,0.25)
Q1_SR.c

# IQR Calulation
Max_IQR_SR.c = Q3_SR.c + 1.5*IQR_SR.c
Max_IQR_SR.c
Min_IQR_SR.c = Q1_SR.c + 1.5*IQR_SR.c

# Outliers calculation
sum(career_bat_new$SR.c > Max_IQR_SR.c)
sum(career_bat_new$SR.c < Min_IQR_SR.c)
```
Overall strike rate is right skewed distribution. The literature states 150% strike rate is becoming very common these days. However, any player with a strike rate above 135% can be considered a decent player. One player who has more than ~210% strike rate.   

#4.0 Multivariate Analysis
```{r}


career_bat_new %>% select(Inns.c:NO_HS.c) %>% ggpairs()
```

In this case study anything above 70% will be considered a heavy correlation attribute pairs and between 30% t0 70% correlation value considered as moderately correlated.Any correlation value below, 30% is loosely correlated. From the above plot it can be said that number of innings played and boundaries scored by players in their IPL career shows very strong correlation (~93%). 














