---
title: "Season_Bowl_Ranking"
author: "Karthik"
date: "1/1/2021"
output: html_document
---


#1.0 Libraries
```{r}
library(corrplot)
library(dplyr)
library(nortest)
library(GPArotation)
library(psych)#Barlett Sphericity Test
library(purrr) # keep() function
library(tidyverse)

```


#2.0 Bowler Ranking
##2.1 Ball df preparation

```{r}

season_ball_num <- season_ball %>% as.data.frame(season_ball) %>% select("Player","Mat.bs","Inns.bs","Overs.bs","Mdns.bs","Runs.bs","Wkts.bs","Ave.bs","Econ.bs","SR.bs","four.bs","five.bs","Ct.bs","St.bs","Season.bs") %>% filter(Season.bs %in% c("2019","2018","2017","2016","2015"))  # Extract last 5 year season information only.


#season_ball_num <- season_ball %>% as.data.frame(season_ball) %>% select("Player":"Season.bs") %>% filter(Season.bs %in% c("2019","2018","2017","2016","2015","2014","2013","2012","2011","2010"))

glimpse(season_ball_num) # Glance dataset.

season_ball_num[,-c(1,15)] <- lapply(season_ball_num[,-c(1,15)],as.numeric) # Convert variable type into numeric.


# Data Wrangling by visualizing individual players performance at once.
season_ball_wider <- pivot_wider(season_ball_num, names_sep = ".", values_fill = 0, names_from = Season.bs ,values_from =c(Mat.bs,Inns.bs,Overs.bs,Mdns.bs,Runs.bs,Wkts.bs,Ave.bs,Econ.bs,SR.bs,four.bs,five.bs,Ct.bs,St.bs )) 

#d <- season_ball_num%>%
#  group_by(Season.bs) %>%
 # mutate(row = row_number()) %>%
  #tidyr::pivot_wider(names_from = Season.bs, 
   #                  names_sep = ".", 
    #                 values_fill = 0,
     #                values_from=c(Mat.bs,Inns.bs,Overs.bs,Mdns.bs,Runs.bs,
      #                             Wkts.bs,Ave.bs,Econ.bs,SR.bs,four.bs,five.bs,
       #                            Ct.bs,St.bs)) %>%
  #select(-row)

season_ball_wider
dim(season_ball_wider)
#season_ball_wider[is.na(season_ball_wider)]<- 0 # Replace NA into 0.
```

##2.2 Missing Values
```{r}
anyNA(season_ball_wider)

```


There are no missing values in the given seasonal batting statistics dataset.

##2.3 Normality Check
Shapiro Wilk's test is applied to check the normality of the attributes.

Ho: Attribute value follows a normal distribution.
Ha: Attribute value does not follow a normal distribution.

```{r}

dim(season_ball_wider)
season_ball_numeric =season_ball_wider %>% keep(is.numeric) # Separate only Numeric variable(i.e. Except "Player" attribute.)


#Shapiro Test
#-------------
#Apply shapiro test to confirm normalization.

f <- function(x) 
  {
  
  if (diff(range(x)) == 0) list() else shapiro.test(x)
  
  } # Function to apply shapiro test

season_ball_shap <- apply(season_ball_numeric[,], 1, f) # Apply shapiro test.
season_ball_shap

```
Shapiro test on attributes failed to follow a normal distribution. Hence, normalization is necessary.


###2.3.1 Scale the data
```{r}
season_ball_scale <- scale(season_ball_numeric,center = TRUE,scale=TRUE) # Scaling
head(season_ball_scale,5)

#season_ball_scale[is.nan(season_ball_scale)]<-0

#season_ball_scale[is.na(season_ball_scale)]<-0
```

# 3.0 Correlation Analysis
##3.1 Corrplot

```{r}
season_ball_scale_edit=season_ball_scale[,-51] # five.bs 2015 contains zero values. 
season_ball_scale_cor <- cor(season_ball_scale_edit) # Generating Correlation Matrix
season_ball_scale_cor
corrplot(season_ball_scale_cor, method= "circle",type="upper") # Plot Correlation Matrix


#summary(season_ball_numeric)
```

#4.0 Barlett Sphericity Test
This test confirms whether we can apply PCA on the given dataset. Hypothesis of the test as follows.

Ho: Data dimension reduction is not possible.
Ha: Data dimension reduction is possible.

```{r}

print(cortest.bartlett(season_ball_scale_cor,nrow(season_ball_scale_edit))) # Bartlett test to confirm possibility of PCA application on the dataset.
```
P-Value is less than 5%. Hence, Ho is failed to accept.Thus, dimentionality reduction is possible.

#5.0 PCA Transformation
## 5.1 Eigen Values
```{r}
A.ball <- eigen(season_ball_scale_cor) # Calculate Eignen Values of the Correlation Matrix
eigenValues.ball <-A.ball$values # Tabulating Eigen Values 
eigenValues.ball
eigenVectors.ball <- A.ball$vectors # Tabulating Eigen Vectors
eigenVectors.ball

```

## 5.2 Plot Scree Plot
```{r}
print(eigenValues.ball,digits=2) # Eignen Values
plot(eigenValues.ball,type="lines",xlab="Principal Components",ylab="Eigen Values")# Plot Eignen Values
```
## 5.3  PCA - Loadings and Communality

```{r}
PCA.ball <-principal(season_ball_scale_edit,nFactors=length(season_ball_scale_edit),rotate="varimax") #Apply PCA with all variables
PCA.ball

```

##5.4  Getting Factor Scores

```{r}
newdf.ball <- PCA.ball$scores
dim(newdf.ball)
dim(season_ball_wider)

newdf.ball = as.data.frame(newdf.ball)
season_ball_rank <- cbind(season_ball_wider$Player,newdf.ball)
range(season_ball_rank$PC1)

```

#6.0 Factor Score Ranking
```{r}
season_ball_rank =season_ball_rank[order(-season_ball_rank$PC1),]
View(season_ball_rank)
season_ball_rank <- rowid_to_column(season_ball_rank, "Bowler_Ranking")
colnames(season_ball_rank)<- c("Bowler_Ranking","Bowler","Factor_bowling_score")

season_ball_rank
```


## 6.1 Cleaning Names
```{r}

season_ball_rank$Bowler <- str_to_lower(season_ball_rank$Bowler) # Converting Player Names into lowercase.

season_ball_rank$Bowler <- gsub(" ","_",season_ball_rank$Bowler)#Replace single space into "_"
season_ball_rank$Bowler <- trimws(gsub("\\s+","",season_ball_rank$Bowler))#Replace double space into none.

season_ball_rank
```






