---
title: "Season_bat"
author: "Karthik"
date: "11/29/2020"
output: html_document
---

# 1.0 Dataset Download
```{r}
season_bat <- season_bat_df # Storing Seasonal Performance dataset under another .df
```


# 2.0 EDA
## 2.1 Sanity Check
### 2.1.2 Variable Names
```{r}
# Chaning Coloumn Names
colnames(season_bat) <- c("Player","Mat.s","Inns.s","NO.s","Runs.s","HS.s",
                        "Ave.s","BF.s","SR.s","hundreds.s","fifties.s","ducks.s",
                        "four.s","six.s","Season.s","NO_HS.s") 

colnames(season_bat) # Display column names.

```
###2.1.3 Variable Class
```{r}
str(season_bat) # Display variable classes

# Transforming variable classes.
season_bat$Player <- as.factor(season_bat$Player)
season_bat$Mat.s <- as.integer(season_bat$Mat.s)
season_bat$Inns.s <- as.integer(season_bat$Inns.s)
season_bat$NO.s <- as.integer(season_bat$NO.s)
season_bat$Runs.s <- as.integer(season_bat$Runs.s)
season_bat$BF.s <- as.integer(season_bat$BF.s)
season_bat$hundreds.s <- as.factor(season_bat$hundreds.s)
season_bat$fifties.s <- as.factor(season_bat$fifties.s)
season_bat$ducks.s <- as.factor(season_bat$ducks.s)
season_bat$four.s <- as.integer(season_bat$four.s)
season_bat$six.s <- as.integer(season_bat$six.s)
season_bat$HS.s <- as.integer(season_bat$HS.s)
season_bat$NO_HS.s <- as.factor(season_bat$NO_HS.s)
season_bat$Season.s <- as.factor(season_bat$Season.s)

# Generating 5 Number summary f all attributes.
summary(season_bat)
```
5 Numer summary is generated for the seasonal dataset attributes.

## 2.2  Missing Values
```{r}
anyNA(season_bat)
```
Data set looks clean. Let us proceed with the EDA.
Selecting the best player of the season.

## 2.3 Variable Transformation.
```{r}

#Transforming Variables
season_bat_new <- season_bat %>% 
  transmute(
    Player,
    Mat.s,
    Inns.s,
    NO.s,
    Runs.s,
    HS.s,
    Ave.s,
    BF.s,
    SR.s,
    hundreds.s,
    fifties.s,
    ducks.s,
    boundaries.s=four.s+six.s,
    Season.s,
    NO_HS.s) %>% 
  select(Player,Inns.s,Runs.s,Ave.s,SR.s,boundaries.s,NO_HS.s)

```
#3.0 Univariate Analysis.

```{r}
# Histogram of Numeric Attributes
season_bat_new %>% keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value))+
  facet_wrap(~key,scales="free")+
  geom_histogram(binwidth=0.5,color="blue",fill="white")
colnames(season_bat)# Displaying Coloumn names.
```
Selected 5 attributes show right skewed distribution.



##3.1 Key performance Indicator.
```{r}
## Most Played
season_bat %>% 
  select(Player,Mat.s) %>%  
  filter(Mat.s==max(Mat.s)) # Most number of matches played by a batsman

## Least Played
season_bat %>% 
  select(Player,Mat.s) %>%  
  filter(Mat.s==min(Mat.s)) %>% 
  count() # All players have played at least one IPL match in their career. 

## Most Batted
season_bat %>% 
  select(Player,Inns.s) %>% 
  filter(Inns.s==max(Inns.s)) 


## Most Batted Players
season_bat %>% 
  select(Player,Inns.s) %>% 
  filter(Inns.s==min(Inns.s)) %>%count()/nrow(season_bat)*100 

# Top 20 IPL Experienced Players
Season_ten_inn <- top_n(season_bat,10,Inns.s) 
Season_ten_inn 

# Top 20 Batting Average 
season_ave <- top_n(season_bat,20,Ave.s) %>% arrange(desc(Ave.s))
season_ave

# Top 20 Strike Rate
season_sr <- top_n(season_bat,20,SR.s) %>% arrange(desc(SR.s))
season_sr

# Top 20 Highest Scored Batsmen
season_HS <- top_n(season_bat,20,HS.s) %>%  arrange(desc(HS.s))
season_HS

```

## 3.2 Best Experienced Seasonal Batsmen
Let us analyse the top most innings played batsmen strike rate information.
```{r}
# Experienced bowlers with Best Economy Rate

Season_ten_inn %>% 
  ggplot(aes(x=reorder(Player,SR.s),y=SR.s,fill = SR.s))+ 
  geom_bar(stat="Identity")+
  coord_flip()+
  scale_y_continuous("Batting Strike Rate")+
  scale_x_discrete("Batsmen Name")
```


#5.0 Multivariate Analysis

##5.1 Correlation Matrix
Strength of the relationship between each attributes.    

```{r}
#Generating Correlation matrix of numeric attributes
corrmatrix <-season_bat_new %>% 
  keep(is.numeric) %>%  
  cor() 
# Approximate the correlation matrix element to 2 decimal places.
round(corrmatrix,2)
# Plot the correlation matrix
corrplot(corrmatrix,method='circle',type='upper',order="hclust")
```

##5.2 GGpair Plot
```{r}

season_bat_new %>% select(Inns.s:NO_HS.s) %>% ggpairs()

colnames(season_bat_new)
```
As expected "Runs" and "boundaries" show strong correlation of 98%. In addition to that, "Runs" & "Inns", "Ave" & "Runs" and "boundaries" & "Ave" show a moderate correlation of 88.3%,74.9% and 72.7% respectively.

An interesting observation is strike rate of both levels of players who have got out and continued the innings with highest score distribution show sign of normal distribution of densed bars at the centre.






