---
title: "H1N1 Flu Vaccine Intake Casestudy"
author: "Karthik"
date: "10/21/2020"
output: html_document
---

# 1.0 WORKING DIRECTORY
```{r}
setwd("F:/Data Science/Casestudies/Trabeya/Driven Data/")

getwd()

```

#2.0 LIBRARIES
```{r}
#library("MASS") 
library("visdat") # Visualising the Missing Values
library("ggplot2") # Data distrbution Visualisations
library("naniar") # Missing Value Visualisation
library("dplyr") # Data Transformation
library("caTools") # Random Split of Dataset
library("tidyverse") 
library("car")
library("rpivotTable") # Attribute Visualization
```

#3.0 IMPORT DATASETS
The Datasource : Data Driven Website
Topic : Casestudy on H1N1 Flu vaccine received rate.
```{r}
# Downloading Data
RawData = read.csv("training_set_features.csv") # Main Dataset for analysis
TestFeatures = read.csv("test_set_features.csv") # Final dataset for Predictions
Train_Labels = read.csv("training_set_labels.csv")# Dataset Lable Information
```

#4.0 DATA PREPARATION

As a first step of the data analysis process, let us conduct an EDA on the given dataset. Hence, this section talks about the sanity test conducted to clean the data./
##4.1 Sanity Check
Given dataset contains 26,707 observations with 36 different attributes.
```{r}

dim(RawData) # Dimensions of the dataset
```

##  4.1.1 Data Cleaning
The table below illustrates the top 10 observations of the dataset. The table labels have used lot of abbreviations. As a first step of the sanity test label names have been changed into meaning full names.Furthermore, all variable classes have been converted into factors using "lapply" function.
```{r}
head(RawData,10)

```

```{r}
colnames(RawData) # Displays coloumn names of "RawData" dataset.

# Merging Train feature & Train Labels dataset.
RawData <- merge(RawData,Train_Labels,by="respondent_id",all=TRUE)

# Creating Another Dataset - Maintaining the original dataset unchanged.
Dataset<-RawData 

# Modifying the header names
names(Dataset) <- c("ID","Concern","Knowledge","Antiviral_Meds","Social_Distancing","Face_Mask","Wash_Hands","Large_Gatherings","Outside_Home","Touch_Face","H1N1_Vac_Recommendation","Seasonal_Vac_Recommendation","Choronic_Medical_History","Six_Months_Child_Contact","Healthcare_Worker","Health_Insurance","Opinion_H1N1_vacc_effective","Opinion_H1N1_Risk","Opinion_H1N1_Sick_from_Vacc","Opinion_Seas_Vacc_Effective","Opinion_seas_risk","Opinion_Seas_Sick_from_Vacc","Age_Group","Education","Race","Sex","Income_Poverty","Marital_Status","Property","Employment_Status","Geography","Census_msa","Adults","Children","Employment_Industry","Occupation","H1N1_Vaccine","Seasonal_Vaccine")

# Converting 1:36 variables classes into Factor.
Dataset[,] <-lapply(Dataset[,],factor)

# Variable class analysis
str(Dataset)

# Output variable names.
colnames(Dataset) 
```
Below "summary commands" indicate the 5 number summary of 36 attributes.However, let us investigate on the missing values of the variables before proceeding with indepth EDA.

```{r}
summary(Dataset)
```
In addition to missing values, it can be seen from the above 5 number summary that, following attributes have untitled factor levels.

Attributes         |Number of Observations
-------------------|---------------------
Education          |1407
Income Poverty     |4423
Martial Status     |1408
Property           |2042
Employment Status  |1463
Employment Industry|13330
Occupation         |13470


Section 4.1.2 below, converts all untitled factor levels as NA entries. This standardizes the missing value visualization and interpretation process.

## 4.1.2 Data Level Cleaning.

```{r}

#Education
levels(Dataset$Education) <- c(NA,"< 12 Years","12 Years","College Graduate","Some College")
levels(Dataset$Income_Poverty) <- c(NA, "<= $75,000,Above Poverty","> $75,000","Below Poverty")
levels(Dataset$Marital_Status) <- c(NA,"Married","Not Married")
levels(Dataset$Property) <- c(NA,"Own","Rent")
levels(Dataset$Employment_Status) <-c(NA,"Employed","Not in Labor Force","Unemployed")
levels(Dataset$Employment_Industry) <-c(NA,"arjwrbjb","atmlpfrs","cfqqtusy","dotnnunm","fcxhlnwr","haxffmxo","ldnlellj","mcubkhph","mfikgejo","msuufmds","nduyfdeo","phxvnwax","pxcmvdjn","qnlwzans","rucpziij","saaquncn","vjjrobsf","wlfvacwt","wxleyezf","xicduogh","xqicxuve")

levels(Dataset$Occupation) <-c(NA,"bxpfxfdn","ccgxvspp","cmhcxjea","dcjcmpih","dlvbwzss","emcorrxb","haliazsg","hfxkjkmi","hodpvpew","kldqjyjy","mxkfnird","oijqvulv","pvmttkik","qxajmpny","rcertsgn","tfqavkke","ukymxvdu","uqqtjvyb","vlluhbov","xgwztkwe","xqwwgdyp","xtkaffoo","xzmlyyjv")
summary(Dataset)

```

## 4.1.3 Missing Value Analysis

```{r}
anyNA(Dataset) # Checking for Missing Values at the dataset
sum(is.na(Dataset))
```
Well, the output of "anyNA" command reveals that there are missing values in the given dataset.60,762 entries have been missing. Hence,an in depth missing value analysis is needed to understand the reason and if any repetitive pattern can be observed.

```{r}
# Tabulating Coloumn wise Missing value sums
colSums(is.na(Dataset))
```
Based on the above output, "Occupation","Employment Industry" and "Health Insurance" are top 3 attributes containing missing values.(ie: 13,470,13,330 & 12,274 respectively). Upcoming chapters discusses a visual representation of missing values.

### 4.1.3.1 Visualising Missing Values
```{r}
# Class wise Missing Value Plot
vis_dat(Dataset,warn_large_data = FALSE)
```
The above output gives a rough indication on the spread of missing values.Also, note that, all 36 attribute have been converted into factor class. The attributes " Employment Industry" & "Occupation"  shows a very close relationship on the missing values.

```{r}
vis_miss(Dataset,warn_large_data=FALSE)# Percentage of missing values
```

Based on the above plot there are 6% of individual feature information are missing from the given dataset. The NA entry distribution indicates an importance of data collection method study. Insights of missing value analysis definitely improves the quality of data collection process.

I would like to discuss little bit about the missing value theory before we commence any further analysis.Following are the types of missing values that can be observed in any given dataset.

1) MCAR (Missing Completely At Random) - Missing values do not contain any noticeable patterns. It might be influenced by external factors. Best way to handle MCAR type missing entries would be completely omitting those observations. But it's recommended to study, because at most cases it tend to reveal insights on data collection.

2) MAR (Missing At Random) - Missing Values are caused due to other attributes in the dataset.

3)MNAR (Missing Not At Random) - Missing Values are caused due to the attribute itself.This event could happen possibly due to error in capturing certain range of measurements or certain types of information.
```{r}
# Plot on Missing Values among all variables
gg_miss_var(Dataset)
```
The above plot have been sliced into one more dimensions based on people who have taken their H1N1 vaccine and not. Majority of the missing values can be observed among the people who have not received their vaccine yet.

```{r}
gg_miss_var(Dataset,facet= H1N1_Vaccine) # Missing Value clustering based on H1N1 Vaccine
order(round(prop.table(table(Dataset$Employment_Industry,Dataset$H1N1_Vaccine)),3))
#round(prop.table(table(Dataset$Employment_Industry,Dataset$H1N1_Vaccine)),2)*100
```
Missing value analysis indicates a clear pattern among "Occupation" and "Employment Industry" attributes. However, both these attributes have used abbreviations as levels so, it's it's not possible to apply particular industry related statistics or knowledge to treat those NA entries. Treating without prior domain knowledge may create biasness or could redirect the analysis process in negative direction.

In addition to that, the third most attribute with NA, "Health Insurance" contributes about 45.96%.This can be treated using "knn imputation" technique. However, as a first level of analysis, I have decided to consider observations with complete rows of information only. 


```{r}
# Extracting observations containging complete information only.
Data = Dataset[complete.cases(Dataset),]
dim(Data) # Dimensions of new complete dataset.
```

# 5.0 Univariate Analysis

This chapter discusses the univariate anlaysis on H1N1 and Seasonal Vaccine received rate among all the observations.


## 5.1 H1N1 Vaccine
```{r}
ggplot(Data) +
 aes(x = H1N1_Vaccine, fill = H1N1_Vaccine, colour = H1N1_Vaccine) +
 geom_bar() +
 scale_fill_viridis_d(option = "viridis") +
 scale_color_viridis_d(option = "viridis") +
 labs(x = "H1N1 Vaccine Received/Not Received", y = "Frequency", title = "H1N1 Vaccination Received Status Spreadness", subtitle = "H1N1 Vaccine Received/Not Received Vs Frequency") +
 theme_minimal()
round(prop.table(table(Data$H1N1_Vaccine)),3)*100
```
## 5.2 Seasonal Vaccine

```{r}
ggplot(Data) +
 aes(x = Seasonal_Vaccine, fill = Seasonal_Vaccine, colour = Seasonal_Vaccine) +
 geom_bar() +
 scale_fill_viridis_d(option = "viridis") +
 scale_color_viridis_d(option = "viridis") +
 labs(x = "Seasonal Vaccine Received/Not Received", y = "Frequency", title = "Seasonal Vaccination Received Status Spreadness", subtitle = "Seasonal Vaccine Received/Not Received Vs Frequency") +
 theme_minimal()

round(prop.table(table(Data$Seasonal_Vaccine)),3)*100
```

In summary it can be said that, majority of the observations have not received any vaccines. However, there are about 30% and 45% among the overall test patients have received H1N1 flu vaccine and Seasonal Vaccine respectively. 

In the upcoming Chapters, I have decided to add an additional dimension to interpret the attributes of data.

#6.0 Multivariate Analysis
Excluding "ID" column and 2 target variables (H1N1 & Seasonal Vaccine), there are about 35 independent variables and can be grouped them into 5 main categories based on their features as listed below.
1) Social
2) Economic
3) Demographic
4) Personal opinions
5) Behavioral

##6.1 Social Features.
Attributes which illustrates the social behaviors of the observant were analyzed to identify hidden patterns in the success rate of receiving H1N1 & Seasonal vaccines. 
The variables are:
1) Social Distancing
2) Large Gatherings
3) Outside Home
4) Education
5) Employment Industry
6) Occupation

###6.1.1 H1N1 Vaccine
```{r}
ggplot(Data) +
 aes(x = Social_Distancing, fill = Education, colour = Education) +
 geom_bar() +
 scale_fill_viridis_d(option = "viridis") +
 scale_color_viridis_d(option = "viridis") +
 labs(x = "Left : Not Received H1N1 , Right: Received H1N1 with Social Distance Practice", y = "Frequency", title = "Multivariate Analysis", subtitle = "Socail Distancing Vs. Education Vs H1N1 Vaccine.") +
 theme_minimal() +
 theme(legend.position = "left") +
 facet_wrap(vars(H1N1_Vaccine), scales = "free_y")
```

```{r}
rpivotTable(Data,rows="Education", cols=c("Seasonal_Vaccine","Social_Distancing"),aggregatorName = "Count as Fraction of Total", rendererName = "Table", width="100%", height="400px")
```


### 6.1.2 Seasonal Vaccine
```{r}
ggplot(Data) +
 aes(x = Social_Distancing, fill = Education, colour = Education) +
 geom_bar() +
 scale_fill_viridis_d(option = "viridis") +
 scale_color_viridis_d(option = "viridis") +
 labs(x = "Left : Not Received Seasonal , Right: Received Seasonal Vaccine with Social Distance Practice", y = "Frequency", title = "Multivariate Analysis", subtitle = "Social Distancing Vs. Education Vs Seasonal Vaccine.") +
 theme_minimal() +
 theme(legend.position = "left") +
 facet_wrap(vars(H1N1_Vaccine), scales = "free_y")
```
```{r}
rpivotTable(Data,rows="Education", cols=c("H1N1_Vaccine","Social_Distancing"),aggregatorName = "Count as Fraction of Total", rendererName = "Table", width="100%", height="400px")
```


```{r}
rpivotTable(Data)
rpivotTable(Data,rows="Education", cols=c("Seasonal_Vaccine","Social_Distancing"),aggregatorName = "Count as Fraction of Total", rendererName = "Table", width="100%", height="400px")
```



##6.2 Economical Feature Variables
The independent variable associated with economical features are as listed below, in order to interpret the hidden pattern related to the receivers/non receiver of both H1N1 and Seasonal vaccine.
1) Health Insurance
2) Income Poverty
3) Property
4) Employment Status

###6.2.1 H1N1 Vaccine
```{r}

ggplot(Data) +
 aes(x = Health_Insurance, fill = Income_Poverty) +
 geom_bar(position = "dodge") +
 scale_fill_viridis_d(option = "viridis") +
 labs(x = "Health Insurance Status", y = "Frequency", title = "MuliVariate Analysis - H1N1 Vaccine", subtitle = "Health Insurance Vs. Income Poverty", fill = "Income Level") +
 theme_minimal() +
 facet_wrap(vars(H1N1_Vaccine))
```

```{r}
rpivotTable(Data,rows="Income_Poverty", cols=c("H1N1_Vaccine","Health_Insurance"),aggregatorName = "Count as Fraction of Total", rendererName = "Table", width="100%", height="400px")
```

###6.2.2 Seasonal Vaccine
```{r}

ggplot(Data) +
 aes(x = Health_Insurance, fill = Income_Poverty) +
 geom_bar(position = "dodge") +
 scale_fill_viridis_d(option = "viridis") +
 labs(x = "Health Insurance Status (0: NO;1:YES)", y = "Frequency", title = "MuliVariate Analysis - Seasonal Vaccine (0: Not Received ; 1: Received)", subtitle = "Health Insurance Vs. Income Poverty", fill = "Income Level") +
 theme_minimal() +
 facet_wrap(vars(Seasonal_Vaccine))
```

```{r}
rpivotTable(Data,rows="Income_Poverty", cols=c("Seasonal_Vaccine","Health_Insurance"),aggregatorName = "Count as Fraction of Total", rendererName = "Table", width="100%", height="400px")
```

##6.3 Demographical Information.
Analyzing information on observance's geographical location for insights in receiving H1N1 and Seasonal vaccines.

1)Geography
2)Census_MSA

###6.3.1 H1N1 Vaccine
```{r}
ggplot(Data) +
 aes(x = Census_msa, fill = Census_msa) +
 geom_bar(position = "dodge") +
 scale_fill_viridis_d(option = "viridis") +
 labs(x = "Census_MSA", y = "Frequency", title = "Multi Variate Analysis - H1N1 Vaccine Geographical Cluster", subtitle = "Census_MSA Vs. H1N1 Vaccine") +
 theme_minimal() +
 facet_wrap(vars(H1N1_Vaccine), scales = "free")
```
```{r}
rpivotTable(Data,rows="Census_msa", cols=c("H1N1_Vaccine"),aggregatorName = "Count as Fraction of Total", rendererName = "Table", width="100%", height="400px")
```


```{r}
ggplot(Data) +
 aes(x = Health_Insurance, fill = Income_Poverty) +
 geom_bar(position = "dodge") +
 scale_fill_viridis_d(option = "viridis") +
 labs(x = "Health Insurance Status", y = "Frequency", title = "MuliVariate Analysis - H1N1 Vaccine", subtitle = "Health Insurance Vs. Income Poverty", fill = "Income Level") +
 theme_minimal() +
 facet_wrap(vars(H1N1_Vaccine))
#esquisser(Data)
```

###6.3.2 Seasonal Vaccine
```{r}
ggplot(Data) +
 aes(x = Census_msa, fill = Census_msa) +
 geom_bar(position = "dodge") +
 scale_fill_viridis_d(option = "viridis") +
 labs(x = "Census_MSA", y = "Frequency", title = "Multi Variate Analysis - Seasonal Vaccine Geographical Cluster", subtitle = "Census_MSA Vs. H1N1 Vaccine") +
 theme_minimal() +
 facet_wrap(vars(Seasonal_Vaccine), scales = "free")
```
```{r}
rpivotTable(Data,rows="Census_msa", cols=c("Seasonal_Vaccine"),aggregatorName = "Count as Fraction of Total", rendererName = "Table", width="100%", height="400px")
```


## 6.4 Obsevants Opinions
Each observant were asked various H1N1 and Seasonal vaccines related questions to rate certain important aspects of this scenario to get better insights on individuals understanding about this flu.

1) Concern
2) Knowledge
3) H1N1_Vaccine_Recommendation
4) Seasonal_Vaccine_Recommendation
5) Opinion_H1N1_vacc_effective
6) Opinion_H1N1_Risk
7) Opinion_H1N1_Sick_from_Vacc 
8) Opinion_Seas_Vacc_Effective
9) Opinion_seas_risk
10)Opinion_Seas_Sick_from_Vacc

### 6.4.1 H1N1 Vaccince
```{r}
ggplot(Data) +
 aes(x = Concern, fill = Opinion_H1N1_vacc_effective) +
 geom_bar(position = "dodge") +
 scale_fill_viridis_d(option = "viridis") +
 labs(x = "Level of Concern about H1N1 flu ( 0: Not Concern, 1: Not Very Concern, 2:Somewhat Cocerned, 3:Very Concerned )", y = "Frequency", title = "Multivariate Analysis - H1N1 Vaccine- Candidates Rating", subtitle = "H1N1 Vaccine Effective Rating Vs. Concern", fill = "H1N1 Vaccine Effectiveness Rating") +
 theme_minimal() +
 facet_wrap(vars(H1N1_Vaccine))
```
```{r}

rpivotTable(Data,rows="Opinion_H1N1_vacc_effective", cols=c("H1N1_Vaccine","Concern"),aggregatorName = "Count as Fraction of Total", rendererName = "Table", width="100%", height="400px")
```


### 6.4.2 Seasonal Vaccine
```{r}
ggplot(Data) +
 aes(x = Concern, fill = Opinion_Seas_Vacc_Effective) +
 geom_bar(position = "dodge") +
 scale_fill_viridis_d(option = "viridis") +
 labs(x = "Level of Concern about H1N1 flu ( 0: Not Concern, 1: Not Very Concern, 2:Somewhat Cocerned, 3:Very Concerned )", y = "Frequency", title = "Multivariate Analysis - H1N1 Vaccine- Candidates Rating", subtitle = "Seasonal Vaccine Effective Rating Vs. Concern", fill = "Seasonal Vaccine Effectiveness Rating") +
 theme_minimal() +
 facet_wrap(vars(H1N1_Vaccine))
```
```{r}

rpivotTable(Data,rows="Opinion_seas_Vacc_Effective", cols=c("Seasonal_Vaccine","Concern"),aggregatorName = "Count as Fraction of Total", rendererName = "Table", width="100%", height="400px")
```

# Behavioural Features Information

Candidates behavioral related information were gathered and grouped for analysis.

1) Face_Mask
2) Wash_Hands
3) Touch_Face
4) Chronic_Medical_History
5) Six_Months_Child_Contact 
6) Healthcare_Worker
7) Age_Group
8) Race
9) Sex
10)Martial Status
11)Adults
12)Children

### 6.4.2 H1N1 Vaccine
```{r}
ggplot(Data) +
 aes(x = Face_Mask, fill = Sex) +
 geom_bar(position = "dodge") +
 scale_fill_viridis_d(option = "viridis") +
 labs(x = "Face Mask Usage ( 0: No Mask ; 1: Yes )", y = "Frequency", title = "Multivariate Analysis - H1N1 Vaccine on Behavioural Cluster", subtitle = "Face Mask Usage Vs. Gender", fill = "Candidate Gender") +
 theme_minimal() +
 facet_wrap(vars(H1N1_Vaccine))
```

```{r}

rpivotTable(Data,rows="Face_Mask", cols=c("H1N1_Vaccine","Sex"),aggregatorName = "Count as Fraction of Total", rendererName = "Table", width="100%", height="400px")
```

### 6.4.3 Seasonal Vaccine
```{r}
ggplot(Data) +
 aes(x = Face_Mask, fill = Sex) +
 geom_bar(position = "dodge") +
 scale_fill_viridis_d(option = "viridis") +
 labs(x = "Face Mask Usage ( 0: No Mask ; 1: Yes )", y = "Frequency", title = "Multivariate Analysis - H1N1 Vaccine on Behavioural Cluster", subtitle = "Face Mask Usage Vs. Gender", fill = "Candidate Gender") +
 theme_minimal() +
 facet_wrap(vars(Seasonal_Vaccine))
```

```{r}

rpivotTable(Data,rows="Face_Mask", cols=c("Seasonal_Vaccine","Sex"),aggregatorName = "Count as Fraction of Total", rendererName = "Table", width="100%", height="400px")
```
The above different analysis have given insights on importance and influencing variables which may come in handy during the model building process. Let us prepare the train and test dataset. 

#5.0 Train & Test Data Split

In order to train and test the model before conducting the actual evaluation on the given test dataset, the observations from the complete dataset were randomly sliced into 70% and 30% proportions using "sample.split" function.
```{r}
set.seed(369) # Assigning random seed value. 
split=sample.split(Data$H1N1_Vaccine,SplitRatio = 0.7)
train.D <- subset(Data,split==TRUE)  # Train Dataset
test.D <- subset(Data ,split==FALSE) # Test Dataset

dim(train.D) # Training Dataset Dimention 
dim(test.D)  # Testing Dataset Dimention
```

# 6.0 Test Feature Converting Names

```{r}
# Modifying the Test Features header names
names(TestFeatures) <- c("ID","Concern","Knowledge","Antiviral_Meds","Social_Distancing","Face_Mask","Wash_Hands","Large_Gatherings","Outside_Home","Touch_Face","H1N1_Vac_Recommendation","Seasonal_Vac_Recommendation","Choronic_Medical_History","Six_Months_Child_Contact","Healthcare_Worker","Health_Insurance","Opinion_H1N1_vacc_effective","Opinion_H1N1_Risk","Opinion_H1N1_Sick_from_Vacc","Opinion_Seas_Vacc_Effective","Opinion_seas_risk","Opinion_Seas_Sick_from_Vacc","Age_Group","Education","Race","Sex","Income_Poverty","Marital_Status","Property","Employment_Status","Geography","Census_msa","Adults","Children","Employment_Industry","Occupation")

#Education
levels(TestFeatures$Education) <- c(NA,"< 12 Years","12 Years","College Graduate","Some College")
levels(TestFeatures$Income_Poverty) <- c(NA, "<= $75,000,Above Poverty","> $75,000","Below Poverty")
levels(TestFeatures$Marital_Status) <- c(NA,"Married","Not Married")
levels(TestFeatures$Property) <- c(NA,"Own","Rent")
levels(TestFeatures$Employment_Status) <-c(NA,"Employed","Not in Labor Force","Unemployed")
levels(TestFeatures$Employment_Industry) <-c(NA,"arjwrbjb","atmlpfrs","cfqqtusy","dotnnunm","fcxhlnwr","haxffmxo","ldnlellj","mcubkhph","mfikgejo","msuufmds","nduyfdeo","phxvnwax","pxcmvdjn","qnlwzans","rucpziij","saaquncn","vjjrobsf","wlfvacwt","wxleyezf","xicduogh","xqicxuve")

levels(TestFeatures$Occupation) <-c(NA,"bxpfxfdn","ccgxvspp","cmhcxjea","dcjcmpih","dlvbwzss","emcorrxb","haliazsg","hfxkjkmi","hodpvpew","kldqjyjy","mxkfnird","oijqvulv","pvmttkik","qxajmpny","rcertsgn","tfqavkke","ukymxvdu","uqqtjvyb","vlluhbov","xgwztkwe","xqwwgdyp","xtkaffoo","xzmlyyjv")

# Converting 1:36 variables classes into Factor.
TestFeatures[,] <-lapply(TestFeatures[,],factor)
```





