---
title: "Bagging"
author: "Karthik"
date: "6/20/2020"
output: html_document
---


BAGGING:
---------

Baggining technique is used to minimise the variance in the dataset. In the case of an outlier, bagging technique will focus on the entire dataset to create new rows which makes the outliers a less of an influential factor.
#1.0  SPLIT for BAGIING
```{r}
bag <- CHD

bag$Gender <- as.factor(bag$Gender)
bag$Education <- as.factor(bag$Education)
bag$Smoking_Status <- as.factor(bag$Smoking_Status)
bag$BP_Medication <- as.factor(bag$BP_Medication)
bag$Prevalent_Stroke <-as.factor(bag$Prevalent_Stroke)
bag$Prevalent_Hypertension <- as.factor(bag$Prevalent_Hypertension)
bag$Diabetes <- as.factor(bag$Diabetes)
bag$CHD_in_10Yrs <-as.factor(bag$CHD_in_10Yrs)

split.bag <- bag
set.seed(369)
split_bag=sample.split(split.bag$CHD_in_10Yrs,SplitRatio = 0.7)
train.bag <- subset(split.bag,split_bag==TRUE)
test.bag <- subset(split.bag,split_bag==FALSE)
dim(test.bag)

```


#2.0 BAGGING Model Building
```{r}
Bagging <- bagging(CHD_in_10Yrs~.,data=train.bag,
                        control=rpart.control(maxdepth=6,minsplit=3)) # Applying Bagging technique

bag.results.df <- data.frame(real=train.bag$CHD_in_10Yrs)
bag.results.df$bag.prob <- predict(Bagging,train.bag,type='prob')[,2] # Prediction made on train dataset

# Optimal Cutoff
InformationValue::optimalCutoff(ifelse(bag.results.df$real=="1",1,0),bag.results.df$bag.prob)

bag.results.df$bag.class <- ifelse(bag.results.df$bag.prob >=0.21,"1","0")
bag.results.df$bag.class = as.factor(bag.results.df$bag.class)


caret::confusionMatrix(bag.results.df$real,bag.results.df$bag.class) # Calculating the Matrix train data
InformationValue::Concordance(ifelse(bag.results.df$real=='1','1','0'),bag.results.df$bag.prob)

```

#3.0 Test Dataset
```{r}
#Creating a new dataset with the results
bag.results.df.test <- data.frame(real=test.NB$CHD_in_10Yrs)

bag.results.df.test$bag.prob.test <-predict(Bagging,test.bag,type='prob')[,2]

```



Creating class on the predicted model.

```{r}
bag.results.df.test$bag.class.test <- as.factor(ifelse(bag.results.df.test$bag.prob.test >=0.5,'1','0'))
```

#4.0 Test Confussion Matrix
```{r}

caret::confusionMatrix(bag.results.df.test$real,bag.results.df.test$bag.class.test)

InformationValue:: Concordance(ifelse(bag.results.df.test$real=='1','1','0'),bag.results.df.test$bag.prob.test)
```
#5.0 ROC

```{r}
pred.bag <- prediction(bag.results.df.test$bag.prob.test,test.bag$CHD_in_10Yrs)
perf.bag <- performance(pred.bag,"tpr","fpr")
plot(perf.bag,col="blue",main="ROC Plot BAGGING - Test Dataset")+abline(0,1,lty=8,col="red")
```
#6.0 AUC
```{r}
test.auc.bag <- performance(pred.bag,"auc")
test.auc.bag<- as.numeric(test.auc.bag@y.values)
round(test.auc.bag*100,2)
```

#7.0 KS Statistics
```{r}
ks_table.bag <-max(attr(perf.bag, "y.values")[[1]] - (attr(perf.bag, "x.values")[[1]]))

ks_table.bag
```