---
title: "CAPSTONE PROJECT - NOTES - 1"
author: "Karthik"
date: "5/10/2020"
output:
  word_document: default
  html_document: default
---
#1.0 INITIAL PREPARATION
##1.1 Working Directory
```{r}
setwd("F:/Data Science/CAPSTONE/") # Set Working Directory
getwd()
```
##1.2 Importing Dataset
```{r}
dataset <- read.csv("Coronary_heart_risk_study.csv")
data <- dataset 
```
##1.3 Importing Libraries
```{r}
library("mice")
library("ggplot2")
library("naniar") # Visualising Missing Values
library("corrplot") # Correlation Analysis
library ("VIM") # Inspect & Visualize the pattern of missing values and analyse.
library("esquisse")
library("dplyr")
#install.packages("mice")
library("tidyverse")
#install.packAges("VIM")
library("fmsb") # Correlation VIF analysis
```
#-----------------------------------------------------------

#2.0  DETAIL REPORT


##2.1 Visual Inspection of Data

```{r}
colnames(data) # Names of Columns
ncol(data) # Number of Columns

dim(data) # Dimentions of dataset.

#1,3,4,7,6,8,9)
```

##2.2 Understanding of Attributes
```{r}
#View(data) # View the dataset in R.
summary(data) # Generating 5 line summary.


names(data) <-c("Gender","Age","Education","Smoking_Status",
               "No_Cigarattes","BP_Medication","Prevalent_Stroke",
               "Prevalent_Hypertension","Diabetes","Total_Choloestrol",
               "Systolic_BP","Diastolic_BP","BMI",
               "Heart_Rate","Glucose","CHD_in_10Yrs")





CHD = data
#View(data) 
levels(data$Diabetes)

```

###2.2.1 Missing Values Analysis

```{r}
# MISSING VALUE COUNTS
#---------------------
anyNA(data) # Checking for missing values in the dataset
sum(is.na(data))# Total number of Missing Values
colSums(is.na(data)) # Total missing values based on categories.
```

###2.2.2 Correlation Analysis
```{r}
# CORRELATION ANALYSIS

str(CHD)

# Converting all attribute class to numeric to apply corrplot

CHD$Gender <- as.numeric(CHD$Gender)
CHD$Smoking_Status <- as.numeric(CHD$Smoking_Status)
CHD$BP_Medication <- as.numeric(CHD$BP_Medication)
CHD$Prevalent_Stroke <- as.numeric(CHD$Prevalent_Stroke)
CHD$Prevalent_Hypertension <- as.numeric(CHD$Prevalent_Hypertension)
CHD$Diabetes <- as.numeric(CHD$Diabetes)
CHD$Age <- as.numeric(CHD$Age)
CHD$Education <- as.numeric(CHD$Education)
CHD$No_Cigarattes <- as.numeric(CHD$No_Cigarattes)
CHD$Total_Choloestrol <- as.numeric(CHD$Total_Choloestrol)
CHD$Heart_Rate <- as.numeric(CHD$Heart_Rate)
CHD$Glucose <- as.numeric(CHD$Glucose)

# Extracting completed rows only and all columns for corrplot analysis
CHD <- CHD[complete.cases(CHD),] # Without NAs.
#----------------------------------------------------------------------------------
xgdata <- CHD
# Correlation Table
Correlation_Table <- cor(CHD[,1:15])
#Writing correlation table into csv file.
write.csv(Correlation_Table,"Correlation_Table.csv")
# Plotting correlation plot
corrplot((cor(CHD[,1:15])),method="circle",type="full",order="hclust")

```

###2.2.3 Converting Data Class

```{r}
attach(CHD)  # Attaching the dataset.
str(CHD)

CHD$Gender <- as.factor(CHD$Gender)
CHD$Education <- as.factor(CHD$Education)
CHD$Smoking_Status <- as.factor(CHD$Smoking_Status)
CHD$BP_Medication <- as.factor(CHD$BP_Medication)
CHD$Prevalent_Stroke <-as.factor(CHD$Prevalent_Stroke)
CHD$Prevalent_Hypertension <- as.factor(CHD$Prevalent_Hypertension)
CHD$Diabetes <- as.factor(CHD$Diabetes)
CHD$CHD_in_10Yrs <-as.factor(CHD$CHD_in_10Yrs)

#Changing Column title from "Gender" to "Gender"
colnames(CHD)[colnames(CHD)=="Gender"] <-"Gender"
colnames(CHD) # Displaying All column names.

# Changing Binary level to String for "Gender"attribute
levels(CHD$Gender) <-c("Female","Male")
levels(CHD$Gender)

# Changing Binary level to String for "CHD_in_10Yrs"attribute
levels(CHD$CHD_in_10Yrs)<- c("Safe","Risk")

# Changing Binary level to String for "Smoking_Status"attribute
levels(CHD$Smoking_Status)<- c("Non-Smoker","Smoker")

# Changing Binary level to String for "BP Meds"attribute
levels(CHD$BP_Medication) <- c("No","Yes")
table(CHD$BP_Medication)
# Changing Binary level to String for "Prevalent_Stroke" attribute
levels(CHD$Prevalent_Stroke) <- c("No","Yes")

# Changing Binary level to String for "Prevalent_Hypertension" attribute
levels(CHD$Prevalent_Hypertension) <- c("No","Yes")

# Changing Binary level to String for "Diabetes" attribute
levels(CHD$Diabetes) <- c("No","Yes")

```
#------------------------------------------------------------

##3.0 EXPLORATORY DATA ANLYSIS
###3.1 UNIVARIATE ANALYSIS

####3.1.1 Gender Patients

```{r}
Gender_Table= table(CHD$Gender) # Genderwise counts on table

Gender_Table # Calling Gender count on table.
```



####3.1.2 Age
```{r}
attach(CHD)
range(Age) # Min & Max of patients Age

#Histogram
ggplot(data,aes(x=Age))+geom_histogram(color="darkblue",fill='lightblue',binwidth=1)+
  geom_vline(aes(xintercept=mean(Age)),color="black",linetype="dashed",size=1)+
  geom_vline(aes(xintercept=median(Age)),color ="red",linetype="dashed",size=1)+
  labs(title="Age Histogram Plot",x="Age", y = "Count")

# Box plot
boxplot(CHD$Age,horizontal = TRUE,col="blue",main = "Patient's Age Distribution", xlab= "Age in Years")
```

```{r}
# Age OUTLIER CALCULATION

attach(CHD) # Attach CHD dataset.

# OUTLIER CALCULATION
mean(Age) # AverAge Age
median(Age) # Age centre value
Age_IQR = IQR(Age)
Age_IQR

#OUTLIERS - Q3
Q3_Age = quantile(Age,0.75)
Q3_Age = unname(Q3_Age)
Q3_Age

#OUTLIERs - Q1
Q1_Age = unname(quantile(Age,0.25))
Q1_Age

# IQR CALCULATION
Max_Age_IQR = Q3_Age+1.5*Age_IQR # Positive Outlier Calculation
Max_Age_IQR
Min_Age_IQR = Q1_Age-1.5*Age_IQR # Negative Outlier Calculation
Min_Age_IQR

# No Of OUTLIERS
sum(Age > Max_Age_IQR)
sum(Age < Min_Age_IQR)
# Print all Outliers
Age[Age > Max_Age_IQR]
Age[Age < Min_Age_IQR]
```


####3.1.3 Education
```{r}
#range(Education) # Min & Max educational history.
table(Education)
#mean(Education) # averAge Educational years
#median(Education) # Centre value of the Educational distribution.

```

####3.1.4 Current Smoker
```{r}
attach(CHD)
#range(Smoking_Status) # Min & Max 
levels(Smoking_Status) # Classification : Smoker & Non-Smoker
table(Smoking_Status)

```

####3.1.5 Cigarattes Per Day

```{r}
range(No_Cigarattes)# Min & Max cigarattes smoked

#Histogram

ggplot(CHD,aes(x=No_Cigarattes))+geom_histogram(color="darkblue",fill='lightblue',binwidth=1)+
  geom_vline(aes(xintercept=mean(No_Cigarattes)),color="black",linetype="dashed",size=1)+
  geom_vline(aes(xintercept=median(No_Cigarattes)),color ="red",linetype="dashed",size=1)+
  labs(title="Cigarrates per day Histogram Plot",x="Cigs per day", y = "Count")


# Box plot

boxplot(CHD$No_Cigarattes,horizontal = TRUE,col="blue",main = "Patient's Smoking Habbbit Distribution", xlab= "Cigarattes per day")

```

```{r}
# OUTLIER CALCULATION
mean(No_Cigarattes) # averAge cigarattes smoked per day.
median(No_Cigarattes) # Centre value number of cigarattes smoked
No_Cigarattes_IQR = IQR(No_Cigarattes)
No_Cigarattes_IQR

#OUTLIERS - Q3
Q3_No_Cigarattes = quantile(No_Cigarattes,0.75)
Q3_No_Cigarattes = unname(Q3_No_Cigarattes)
Q3_No_Cigarattes

#OUTLIERs - Q1
Q1_No_Cigarattes = unname(quantile(No_Cigarattes,0.25))
Q1_No_Cigarattes

# IQR CALCULATION
Max_No_Cigarattes_IQR = Q3_No_Cigarattes+1.5*No_Cigarattes_IQR # Positive Outlier Calculation
Max_No_Cigarattes_IQR
Min_No_Cigarattes_IQR = Q1_No_Cigarattes-1.5*No_Cigarattes_IQR # Negative Outlier Calculation
Min_No_Cigarattes_IQR

# No Of OUTLIERS
sum(No_Cigarattes > Max_No_Cigarattes_IQR)
sum(No_Cigarattes < Min_No_Cigarattes_IQR)
# Print all Outliers
No_Cigarattes[No_Cigarattes > Max_No_Cigarattes_IQR]
No_Cigarattes[No_Cigarattes < Min_No_Cigarattes_IQR]
```

####3.1.6 BP Medication

```{r}
table(BP_Medication) # Number of patients under BP medication.
```

####3.1.7 Prevalent Stroke

```{r}
table(Prevalent_Stroke) # Number of patients with stroke history.
```

####3.1.8 Prevalent Hypertension
```{r}
table(Prevalent_Hypertension)# Number of patients with Hypertensive history.
```

#
####3.1.9 Diabetes

```{r}
table(Diabetes)# Number of patients with Diabetes history.
```

####3.1.10 Total Choloestrol

```{r}
range(Total_Choloestrol) # Min & Max of tot choloestral level

#Histogram
ggplot(CHD,aes(x=Total_Choloestrol))+geom_histogram(color="darkblue",fill='lightblue',bandwidth=1)+
  geom_vline(aes(xintercept=mean(Total_Choloestrol)),color="black",linetype="dashed",size=1)+
  geom_vline(aes(xintercept=median(Total_Choloestrol)),color ="red",linetype="dashed",size=1)+
  labs(title="Patient's choloestrol Histogram Plot",x="Total Choloestrol", y = "Count")

# Box plot
boxplot(CHD$Total_Choloestrol,horizontal = TRUE,col="red",main = "Patient's choloestrol Distribution", xlab= "Total Choloestrol")
```

```{r}
# OUTLIER CALCULATION
mean(Total_Choloestrol) # AverAge of tot choloestrol
median(Total_Choloestrol) # Centre value of total choloestrol distribution
Total_Choloestrol_IQR = IQR(Total_Choloestrol)
Total_Choloestrol_IQR

#OUTLIERS - Q3
Q3_Total_Choloestrol = quantile(Total_Choloestrol,0.75)
Q3_Total_Choloestrol = unname(Q3_Total_Choloestrol)
Q3_Total_Choloestrol

#OUTLIERs - Q1
Q1_Total_Choloestrol = unname(quantile(Total_Choloestrol,0.25))
Q1_Total_Choloestrol

# IQR CALCULATION
Max_Total_Choloestrol_IQR = Q3_Total_Choloestrol+1.5*Total_Choloestrol_IQR # Positive Outlier Calculation
Max_Total_Choloestrol_IQR
Min_Total_Choloestrol_IQR = Q1_Total_Choloestrol-1.5*Total_Choloestrol_IQR # Negative Outlier Calculation
Min_Total_Choloestrol_IQR

# No Of OUTLIERS
sum(Total_Choloestrol > Max_Total_Choloestrol_IQR)
sum(Total_Choloestrol < Min_Total_Choloestrol_IQR)
# Print all Outliers
Total_Choloestrol[Total_Choloestrol > Max_Total_Choloestrol_IQR]
Total_Choloestrol[Total_Choloestrol < Min_Total_Choloestrol_IQR]
```


####3.1.11 Systolic Blood Pressure

```{r}
range(Systolic_BP) # Min & Max value of Systolic BP.

#Histogram
ggplot(CHD,aes(x=Systolic_BP))+geom_histogram(color="darkblue",fill='lightblue',
                                        bandwidth=1)+
  geom_vline(aes(xintercept=mean(Systolic_BP)),color="black",linetype="dashed",size=1)+
  geom_vline(aes(xintercept=median(Systolic_BP)),color ="red",linetype="dashed",size=1)+
  labs(title="Patient's Systolic BP Histogram Plot",x="Total Systolic BP", y = "Count")

# Box plot

boxplot(CHD$Systolic_BP,horizontal = TRUE,col="red",main = "Patient's Systolic BP", xlab= "Systolic BP")
```



```{r}

# OUTLIER CALCULATION
attach(CHD)
mean(Systolic_BP) # AverAge systolic BP.
median(Systolic_BP) # Centre value of Systolic BP.
Systolic_BP_IQR = IQR(Systolic_BP)
Systolic_BP_IQR

#OUTLIERS - Q3
Q3_Systolic_BP = quantile(Systolic_BP,0.75)
Q3_Systolic_BP = unname(Q3_Systolic_BP)
Q3_Systolic_BP

#OUTLIERs - Q1
Q1_Systolic_BP = unname(quantile(Systolic_BP,0.25))
Q1_Systolic_BP

# IQR CALCULATION
Max_Systolic_BP_IQR = Q3_Systolic_BP+1.5*Systolic_BP_IQR # Positive Outlier Calculation
Max_Systolic_BP_IQR
Min_Systolic_BP_IQR = Q1_Systolic_BP-1.5*Systolic_BP_IQR # Negative Outlier Calculation
Min_Systolic_BP_IQR

# No Of OUTLIERS
sum(Systolic_BP > Max_Systolic_BP_IQR)
sum(Systolic_BP < Min_Systolic_BP_IQR)
# Print all Outliers
Systolic_BP[Systolic_BP > Max_Systolic_BP_IQR]
Systolic_BP[Systolic_BP < Min_Systolic_BP_IQR]

```
####3.1.12 Diastolic Blood Pressure
```{r}
range(Diastolic_BP) # Min & Max of diastolic blood pressure

#Histogram
ggplot(CHD,aes(x=Diastolic_BP))+geom_histogram(color="darkblue",fill='lightblue',bandwidth=1,bin=90)+
  geom_vline(aes(xintercept=mean(Diastolic_BP)),color="black",linetype="dashed",size=1)+
  geom_vline(aes(xintercept=median(Diastolic_BP)),color ="red",linetype="dashed",size=1)+
  labs(title="Patient's Diastolic BP Histogram Plot",x="Diastolic BP", y = "Count")

# Box plot

boxplot(CHD$Diastolic_BP,horizontal = TRUE,col="GREEN",main = "Patient's Diastolic BP Distribution", xlab= "Diastolic BP")

```

```{r}
# OUTLIER CALCULATION
mean(Diastolic_BP) # Average Diastolic_BP
median(Diastolic_BP) # Centre value of Diastolic_BP
Diastolic_BP_IQR = IQR(Diastolic_BP)
Diastolic_BP_IQR

#OUTLIERS - Q3
Q3_Diastolic_BP = quantile(Diastolic_BP,0.75)
Q3_Diastolic_BP = unname(Q3_Diastolic_BP)
Q3_Diastolic_BP

#OUTLIERs - Q1
Q1_Diastolic_BP = unname(quantile(Diastolic_BP,0.25))
Q1_Diastolic_BP

# IQR CALCULATION
Max_Diastolic_BP_IQR = Q3_Diastolic_BP+1.5*Diastolic_BP_IQR # Positive Outlier Calculation
Max_Diastolic_BP_IQR
Min_Diastolic_BP_IQR = Q1_Diastolic_BP-1.5*Diastolic_BP_IQR # Negative Outlier Calculation
Min_Diastolic_BP_IQR

# No Of OUTLIERS
sum(Diastolic_BP > Max_Diastolic_BP_IQR)
sum(Diastolic_BP < Min_Diastolic_BP_IQR)
# Print all Outliers
Diastolic_BP[Diastolic_BP > Max_Diastolic_BP_IQR]
Diastolic_BP[Diastolic_BP < Min_Diastolic_BP_IQR]
```

####3.1.13 BMI
```{r}
range(BMI) # Min & Max of BMI

#Histogram
ggplot(CHD,aes(x=BMI))+geom_histogram(color="darkblue",fill='lightblue',bandwidth=1)+
  geom_vline(aes(xintercept=mean(BMI)),color="black",linetype="dashed",size=1)+
  geom_vline(aes(xintercept=median(BMI)),color ="red",linetype="dashed",size=1)+
  labs(title="Patient's BMI Histogram Plot",x="Body Mass Index", y = "Count")

# Box plot

boxplot(CHD$BMI,horizontal = TRUE,col="purple",main = "Patient's BMI Distribution", xlab= "Body Mass Index")
```


```{r}

# OUTLIER CALCULATION
mean(BMI) # AverAge BMI
median(BMI) # Centre value of BMI
BMI_IQR = IQR(BMI)
BMI_IQR

#OUTLIERS - Q3
Q3_BMI = quantile(BMI,0.75)
Q3_BMI = unname(Q3_BMI)
Q3_BMI

#OUTLIERs - Q1
Q1_BMI = unname(quantile(BMI,0.25))
Q1_BMI

# IQR CALCULATION
Max_BMI_IQR = Q3_BMI+1.5*BMI_IQR # Positive Outlier Calculation
Max_BMI_IQR
Min_BMI_IQR = Q1_BMI-1.5*BMI_IQR # Negative Outlier Calculation
Min_BMI_IQR

# No Of OUTLIERS
sum(BMI > Max_BMI_IQR)
sum(BMI < Min_BMI_IQR)
# Print all Outliers
BMI[BMI > Max_BMI_IQR]
BMI[BMI < Min_BMI_IQR]

```


####3.1.14 Heart Rate

```{r}
range(Heart_Rate) # Min & Max of heart rate.

#Histogram
ggplot(CHD,aes(x=Heart_Rate))+geom_histogram(color="darkblue",fill='lightblue',bandwidth=1)+
  geom_vline(aes(xintercept=mean(Heart_Rate)),color="black",linetype="dashed",size=1)+
  geom_vline(aes(xintercept=median(Heart_Rate)),color ="red",linetype="dashed",size=1)+
  labs(title="Patient's Heart Rate Histogram Plot",x="Heart Rate", y = "Count")

# Box plot

boxplot(CHD$Heart_Rate,horizontal = TRUE,col="purple",main = "Patient's Heart Rate Distribution", xlab= "Heart Rate")
```


```{r}

# OUTLIER CALCULATION
mean(Heart_Rate)  # AverAge Heart Rate
median(Heart_Rate) # Centre value of the Heart Rate distribution.
Heart_Rate_IQR = IQR(Heart_Rate)
Heart_Rate_IQR

#OUTLIERS - Q3
Q3_Heart_Rate = quantile(Heart_Rate,0.75)
Q3_Heart_Rate = unname(Q3_Heart_Rate)
Q3_Heart_Rate 

#OUTLIERs - Q1
Q1_Heart_Rate = unname(quantile(Heart_Rate,0.25))
Q1_Heart_Rate

# IQR CALCULATION
Max_Heart_Rate_IQR = Q3_Heart_Rate+1.5*Heart_Rate_IQR # Positive Outlier Calculation
Max_Heart_Rate_IQR
Min_Heart_Rate_IQR = Q1_Heart_Rate-1.5*Heart_Rate_IQR # Negative Outlier Calculation
Min_Heart_Rate_IQR

# No Of OUTLIERS
sum(Heart_Rate > Max_Heart_Rate_IQR)
sum(Heart_Rate < Min_Heart_Rate_IQR)
# Print all Outliers
Heart_Rate[Heart_Rate > Max_Heart_Rate_IQR]
Heart_Rate[Heart_Rate < Min_Heart_Rate_IQR]
Heart_Rate.txt <-Heart_Rate[Heart_Rate > Max_Heart_Rate_IQR]

write.csv(Heart_Rate,"heart rate Outliers.csv") # Writing outliers into CSV.
```

####3.1.15 Glucose Level

```{r}
range(Glucose) # Min & Max of Glucose level.

#Histogram
ggplot(CHD,aes(x=Glucose))+geom_histogram(color="darkblue",fill='lightblue',bin=60,binwidth = 1)+
  geom_vline(aes(xintercept=mean(Glucose)),color="black",linetype="dashed",size=1)+
  geom_vline(aes(xintercept=median(Glucose)),color ="red",linetype="dashed",size=1)+
  labs(title="Patient's Glucose Level Histogram Plot",x="Glucose Level", y = "Count")


# Box plot

boxplot(CHD$Glucose,horizontal = TRUE,col="purple",main = "Patient's Glucose Level Distribution", xlab= "Glucose Level")
```



```{r}

# OUTLIER CALCULATION
mean(Glucose) # AverAge Glucose level
median(Glucose) # Centre value of Glucose distribution.
Glucose_IQR = IQR(Glucose)
Glucose_IQR 

#OUTLIERS - Q3
Q3_Glucose = quantile(Glucose,0.75)
Q3_Glucose = unname(Q3_Glucose)
Q3_Glucose 

#OUTLIERs - Q1
Q1_Glucose = unname(quantile(Glucose,0.25))
Q1_Glucose

# IQR CALCULATION
Max_Glucose_IQR = Q3_Glucose+1.5*Glucose_IQR # Positive Outlier Calculation
Max_Glucose_IQR
Min_Glucose_IQR = Q1_Glucose-1.5*Glucose_IQR # Negative Outlier Calculation
Min_Glucose_IQR

# No Of OUTLIERS
sum(Glucose > Max_Glucose_IQR)
sum(Glucose < Min_Glucose_IQR)
# Print all Outliers
Glucose[Glucose > Max_Glucose_IQR]
Glucose[Glucose < Min_Glucose_IQR]

```

####3.1.16 CHD Risk Patients

```{r}
table(CHD_in_10Yrs) # Risk & Safe Patients counts 

```

#--------------------------------------------------------------
## 3.2 BI-VARIATE  ANALYSIS
###3.2.1 Demographic Analysis - Gender Vs Age Vs.Education on CHD Risk


```{r}
# Bar Plot - Gender Vs Age Vs. Education on CHD Risk
ggplot(CHD) +
 aes(x = Gender, y = Age, fill = Education) +
 geom_boxplot() +
 scale_fill_viridis_d(option = "plasma") +
 labs(x = "Risk of CHD against Gender", y = "Age (Years)", title = "Box Plot - Demographic Analysis", subtitle = "Age Vs. Gender Vs. Education on CHD Risk", fill = "Years of Education") +
 theme_minimal() +
 facet_wrap(vars(CHD_in_10Yrs))

# Filtering Risk Female through Demographic Analysis.
Risk.Female.df<-CHD %>% select(CHD_in_10Yrs,Gender,Education,Age) %>% filter(CHD_in_10Yrs=="Risk" & Gender =="Female")
head(Risk.Female.df,10)

# Filtering Risk Genders through Demographic Analysis.
Risk.Gender.df<-CHD %>% select(CHD_in_10Yrs,Gender,Education,Age) %>% filter(CHD_in_10Yrs=="Risk" & Gender =="Gender")
Risk.Gender.df

# Filtering Safe Gender through Demographic Analysis.
Safe.Gender.df<-CHD %>% select(CHD_in_10Yrs,Gender,Education,Age) %>% filter(CHD_in_10Yrs=="Safe" & Gender =="Gender")
Safe.Gender.df

# Filtering Safe Female through Demographic Analysis.
Safe.Female.df<-CHD %>% select(CHD_in_10Yrs,Gender,Education,Age) %>% filter(CHD_in_10Yrs=="Safe" & Gender =="Gender")
Safe.Female.df

# Mean summary on Risk Female Age based on Educational history.
by(Risk.Female.df$Age,INDICES=Risk.Female.df$Education,FUN = mean)

# Mean summary on Risk Gender Age based on Educational history.
by(Risk.Gender.df$Age,INDICES=Risk.Gender.df$Education,FUN = mean)

# Mean summary on Safe Female Age based on Educational history.
by(Safe.Female.df$Age,INDICES=Safe.Female.df$Education,FUN = mean)

# Mean summary on Safe Gender Age based on Educational history.
by(Safe.Gender.df$Age,INDICES=Safe.Gender.df$Education,FUN = mean)

```


###3.2.2 Behavioural Analysis - Smoker Vs # Cigs per day on CHD Risk

```{r}
# Bar Plot - Smoker Vs # Cigs per day on CHD Risk

ggplot(CHD) +
 aes(x = Smoking_Status, y = No_Cigarattes, fill = CHD_in_10Yrs) +
 geom_boxplot() +
 scale_fill_hue() +
 labs(x = "Smoking Status", y = "# of Ciggaratte Per Day", title = "Box Plot - Behavioural Analysis", subtitle = "Smoker Vs # of Cigs Per day on CHD", caption = "asd", fill = "Risk on CHD") +
 theme_classic()
```


#### 3.2.2.1 CHD Risk on Non and Smokers.
```{r}
# Filtering Non Smokers Risk on CHD

NonSmoke_CHD.df<-(CHD %>% select(Smoking_Status,CHD_in_10Yrs) %>% filter(Smoking_Status=="Non-Smoker"))
head(NonSmoke_CHD.df,10)
table(NonSmoke_CHD.df) # Table of Risky non smokers

# Filtering Smokers on CHD
Smoke_CHD.df<-(CHD %>% select(Smoking_Status,CHD_in_10Yrs) %>% filter(Smoking_Status=="Smoker"))
table(Smoke_CHD.df)

# Bar plot of Behavioral Analysis of Smokers and Non-Smokers of CHD risk.
ggplot(CHD) +
 aes(x = Smoking_Status, fill = Smoking_Status) +
 geom_bar() +
 scale_fill_brewer(palette = "Set1") +
 labs(x = "Patients Smoking Habbit", 
      y = "Number of Test Patients", 
      title = "Behavioural Analysis", 
      subtitle = "Risk of CHD on Smokers Vs. Non-Smokers", 
      fill = "Behaviour") +
 theme_minimal() +
 facet_wrap(vars(CHD_in_10Yrs))

```


###3.2.3 Health Analysis - Health Condition on CHD Risk
```{r}
# Filtering Patients with clean medical history.
HealthCHD.df <- CHD %>%select(BP_Medication,Prevalent_Hypertension,Prevalent_Stroke,Diabetes,CHD_in_10Yrs) %>% filter(BP_Medication=="No",Prevalent_Hypertension=="No",Prevalent_Stroke=="No",Diabetes=="No") 


table(HealthCHD.df$CHD_in_10Yrs) # Table on patients clean medical history.
#View (CHD) # View the dataset.
write.csv(CHD,"Coronary Heart Disease Analysis.csv")

# Bar plot on Safe & Risk Prediction on people with healthy lifestyle.
ggplot(HealthCHD.df) +
 aes(x = CHD_in_10Yrs, fill = CHD_in_10Yrs) +
 geom_bar() +
 scale_fill_viridis_d(option = "viridis") +
 labs(x = "Safe & Risk Patients", y = "Number of Patients", title = "Health Analysis", subtitle = "Safe & Risk on CHD among Healthy Patients", caption = "No suspision of helath record") +
 theme_minimal()

```

###3.2.3.1 Health Analysis - Diastolic & Systolic BP.
```{r}
# Risk analysis on Systolic BP
Risk.Sys.Dia.df <- CHD %>% select(Systolic_BP,Diastolic_BP,CHD_in_10Yrs) 
range(Risk.Sys.Dia.df$Systolic_BP)

head(Risk.Sys.Dia.df,10)

# Classification of Hypertensive based on Systolic BP and Diastolic BP
Hyp.Class <- mutate(Risk.Sys.Dia.df,Hyp.Class=ifelse(Systolic_BP<120 & Diastolic_BP<80,"Normal",ifelse(Systolic_BP>=120 & Systolic_BP<=139& Diastolic_BP>=80 & Diastolic_BP<=90,"PreHypertension",ifelse(Systolic_BP>=140 & Systolic_BP<160 & Diastolic_BP>=90 & Diastolic_BP <=99,"Hyp StAge-1",ifelse(Systolic_BP>=160 & Diastolic_BP >= 100,"Hyp StAge -2","NO CLASS")))))
head(Hyp.Class,10)

# Bar plot -  Hypertension Classification on CHD Risk
ggplot(Hyp.Class) +
 aes(x = Hyp.Class, fill = Hyp.Class) +
 geom_bar() +
 scale_fill_viridis_d(option = "viridis") +
 labs(x = "Hypertension Classes", y = "Number of Patients", title = "Health Analysis", subtitle = "Hypertension Class Vs CHD Risk", caption = "Classified based WHO recommendation", fill = "Hypertension Classes ") +
 theme_minimal() +
 facet_wrap(vars(CHD_in_10Yrs))
```

```{r}
write.csv(CHD,"Coronary Heart Disease- Modified.csv")

```


